Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-03-30 apj.tex
Remove zero-color discussion

\begin{abstract}
An empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence.
The color changes allowed by the model are almost parallel to the color changes allowed by the dust model of  \citet{1999PASP..111...63F} with floating
total-to-selective extinction parameter $R^F$.  The prior of no parameter covariance used in the training, and the deduced extinctions and color excess
are both qualitatively consistent with $R^F \sim 2.4$.
The derived dust curve is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Inspired that our  two-parameter model, with no conditioning that would steer it toward a dust model, produces behavior similar to that expected from Galactic dust, we extend  the model to three color parameters.
In this new case we obtain consistent results for
all the parameters in common,
plus a detection of a third parameter at
%--- table.py
$(1-5\times10^{-4})$
%-----
confidence;
this new parameter represents supernova color diversity
beyond that tracked by the spectral features and two-parameter dust models.
%A direct incorporation of dust models into our formalism is impossible. An indirect incorporation of dust
%yields in constant-$R^F$ model $R^F \sim 3$, but the implementation is sensitive to systematic uncertainties that arise
%from the need to specify the colors of an unextincted supernova and its use of absolute extinction, neither of which is used in our original model.
\end{abstract}

@@

\subsection{The Two Extrinsic Per-Supernova Parameters $k_0$ and $k_1$}
\color{blue}
The two sets of  $\gamma^i$ parameters  are significantly non-zero. 
None of the 20000 links of 
our Monte Carlo chains for $\gamma$ extend to 0 (see Figures~\ref{global1:fig}--\ref{global5:fig}), we thus claim detection of the
influence of $k_0$ and $k_1$  on supernova magnitudes
with probability $1-5\times 10^{-5}$.

There is a pair of parameter values,  $k_0^0$ and $k_1^0$, that correspond to supernovae experiencing no extrinsic effect.
\color{orange}
For observer filters $X$ and $Y$, we define $A_X(k_0,k_1) =  \gamma^0_X k_0 + \gamma^1_X k_1$, $E(X-Y)=A_X-A_Y$.
The true extrinsic
extinction is
$A^T_X(k_0,k_1)= A_X(k_0,k_1) -  A_X(k^0_0,k^0_1)$.
\color{blue}
The true extrinsic color excess is $E^T(B-V) = A^T_B-A^T_V$, so that the total-to-selective extinction is $R^T_V=A^T_V/E^T(B-V)$.
Our $\langle k_0 \rangle=0$ and  $\langle k_1 \rangle=0$ constraints
do not determine $k_0^0$ and $k_1^0$, and so the extrinsic extinction and color excess per supernova are not products of our analysis.

%our analysis does not specify their values.
%Recall that there are degeneracies between the intrinsic (Eqn.~\ref{ewsiv:eqn}) and
%extrinsic (Eqn.~\ref{dust:eqn}) parameters of our model.  Part of that degeneracy is
%broken by imposing $\langle k_0 \rangle=0$ and  $\langle k_1 \rangle=0$,
%but this choice does not specify the parameter values.

Nevertheless, we are able to examine aggregate properties of the extinction.  The extrinsic extinction and color relation can be
rearranged as
\begin{equation}
A_X(k_0,k_1) = R^T_V \left( A_B(k_0,k_1) -  A_V(k_0,k_1)\right) -R^T_V \left(A_B(k^0_0,k^0_1) + A_V(k^0_0,k^0_1)\right)+  A_X(k^0_0,k^0_1) .
\end{equation}
Making the assumption that $R^T_V$ is the same for all supernovae,
its value can be determined as the slope of the linear relation between $A_X$ and $E(B-V)$.

Figure~\ref{avebv:fig} shows the relationship between $A_V$ and $E(B-V)$.
The full set of supernovae does not coincide with the expectations of a single linear relation; rather the supernova population is peaked
on a ridge line, with a tail distributed below.  That is, for a fixed color excess, the amount of $V$-band extinction is bounded above
but can have values that extend toward lower values.
The assumption of a constant $R^T_V$ is not altogether correct; still we perform a linear fit that gives a slope of
%----- fitz11.py
$R^T_V=2.77 \pm 0.08$,
%----
where the uncertainty is the $68$\% credible interval and does not include model error.

%\begin{comment}
%\color{blue}
%Our model alone cannot determine $R^T_V = A^T_V/E^T(B-V)$ for individual objects, as $k_0^0$ and $k_1^0$ are unconstrained.
%For a qualitative view, we choose zeros  such that the $V$-band extinction and color-excess of the bluest supernova roughly correspond to $E(B-V)=0$,
%$A_V=0$, and such that the ridge line 
%has close to constant $R^T_V$.
%\color{black}
%With this ansatz, we
%estimate $R^T_V$ per supernova shown in Figure~\ref{avebv:fig}, with the expectations of the
% \citet{1999PASP..111...63F} (hereafter referred to as \citetalias{1999PASP..111...63F})  model for fixed $R^F_V$ overplotted.
%With this choice of zeropoints, the ridge line corresponds to $R_V^F\lesssim 2.5$, and even lower values of $R_V^F$ below the ridge line.
%We emphasize that the shape of the locus in this plot (though not the relative positions of the points)  and the flatness
%of the ridge line is specific to our choice of zeropoints, and that our model does not specify the zeropoints.
%A specification of per-supernova $R_V$ requires a model that can identify supernovae that suffer no extrinsic effects,
%for example by modeling  a constraining distribution for the extrinsic parameters \citep{2016arXiv160904470M}.
%\color{black}
%
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.2in]{output11/avebv.pdf}
   %\includegraphics[width=2.8in]{output11/rv.pdf}
   \caption{
   $A_V$ versus $E(B-V)$, and the slope $R_V$ of the linear fit to the data.
   %Right: Shown for qualitative purposes,
   %$R^T_V=A^T_V/E^T(B-V)$ 
   %versus $E^T(B-V)$ for the supernova sample, given an ansatz for the definition of an intrinsic supernova. 
   %Overplotted is the mean and standard deviation of those points in equally divided bins.
   Overplotted are the expected curves for the \citetalias{1999PASP..111...63F} model
   for specific choices of $R^F$. Note that these lines are not straight, which is apparent when comparing to the straight black $R_V=2.78$ line.
   \label{avebv:fig}}
\end{figure}
%\end{comment}

%Until now we have considered $k_0$ and $k_1$ without association with any specific dust model, but they could be interpreted as
%parameters of a single  dust model.
%\color{red}
%Our model for broadband photometry does not specify how spectral features and color parameters affect the supernova spectrum, information
%necessary to predict the magnitude extinction of dust models.
%\color{blue}
%The lack of correspondence between dust models and broadband extinction is quantified when we apply an $R^F_V=2.5$ SN-frame dust to the
%observed spectra used in our sample, to yield a distribution in the quantity $R_V=A_V/(A_B-A_V)$ with mean and standard deviation of $3.11 \pm 0.35$.
%While an accurate study of fundamental dust properties would incorporate the dust law within the model itself, we now proceed nonetheless
%in making a connection between our model outputs and dust.

We now compare our model and its color-coefficient parameters $\gamma^0$ and $\gamma^1$ with the expectations of  dust models.
The 2-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V)
\label{f99:eqn}
\end{equation}
approximately describes dust models.
For the case of
$R^F=2.5$ and $A^F_V=0.1$,
and the measurement of the SALT2
\citet{2007A&A...466...11G} $s=1$, $x_1=0$ SN~Ia template at $B$-band peak, the \citetalias{1999PASP..111...63F} model
gives
$a(U,B,V,R,I)=\{0.96,   1.00,   1.00,   0.97,   0.77\}$ and $b(U,B,V,R,I)=\{  1.77,   0.98,   0.12,  -0.50,  -0.53\}$.
The $F$ superscript is added to distinguish \citetalias{1999PASP..111...63F} parameters.
Over the ranges of
 $R^F$ and $A^F_V$,
and the wavelengths under consideration,  the values of the elements of $a$ and $b$ change but only by $<5$\%.

@@

As a check of the above interpretation, we generate a simulated dataset with the same numbers and uncertainties as our data, based on a
SALT2 $x_1=0$, $c=0$ template and  \citetalias{1999PASP..111...63F} dust.  The supernovae
are assigned $E^F(B-V)$ parameters equal to the $B-V$ color excesses determined in our sample, and $R^F$ values drawn from a log-normal
distribution with a mean of $\ln{2.4}$ and 0.05 standard deviation.  The $\alpha$, $\beta$, and $\eta$ parameters are set to zero, so that dust and
measurement uncertainty are the only sources of variation in magnitudes.  
From the Monte Carlo chains, we find a 68\% credible interval for $\kappa^{-1}_1$ of 2.18--2.40.
As with the analysis of our data, we find that the
$\gamma$'s and \citetalias{1999PASP..111...63F}  vectors for nearly parallel planes, with the $\gamma^0$ vector almost consistent with
$\kappa_1^{-1}=2.4$, as expected given the mean of $\ln{2.4}$ of the input log-normal distribution.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.95in]{output11_sim/plane0.pdf}
   \includegraphics[width=2.95in]{output11_sim/plane1.pdf}
   \caption{
   Two perspectives of $UVI$- spaces
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
   and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines), for a simulated dataset with
   input $R^F$ drawn from a log-normal distribution  with a mean of $\ln{2.4}$ and 0.05 standard deviation.
   The combinations $a(X)+b(X)/2.3$ and  $a(X)+b(X)/2.4$ are shown in the dotted
   lines, the former is 
   almost  perfectly  superimposed on  $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$.
   \label{plane_sim:fig}}
\end{figure}

@@

Another comparison between our model predictions with those of dust is presented in
Figure~\ref{avebv:fig}, which shows the expectations for measured $A(V)$, $E(B-V)$ for $R^F=1.1, 1.9, 2.5, 3.1$.
Note that for our broadband measurements, the relationship between $A(V)$ and $E(B-V)$ is subtly
non-linear.
Qualitatively, the $R^F \sim 2.4$  most closely matches our linear model fit.

>>>>

\begin{abstract}
An empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and the color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence.
The color changes allowed by the model are almost parallel to the color changes allowed by the dust model of  \citet{1999PASP..111...63F} with floating
total-to-selective extinction parameter $R^F$.  The prior of no parameter covariance used in the training, and the deduced color excess
is consistent with $R^F = 2.44$.
The derived dust curve is able to reproduce the optical colors of the highly reddened SN~2014J better than the \citet{1999PASP..111...63F} dust model.
Inspired that our  two-parameter model, with no conditioning that would steer it toward a dust model, produces behavior similar to that expected from Galactic dust, we extend  the model to three color parameters.
In this new case we obtain consistent results for
all the parameters in common,
plus a detection of a third parameter at
%--- table.py
$(1-5\times10^{-4})$
%-----
confidence;
this new parameter represents supernova color diversity
beyond that tracked by the spectral features and two-parameter dust models.
%A direct incorporation of dust models into our formalism is impossible. An indirect incorporation of dust
%yields in constant-$R^F$ model $R^F \sim 3$, but the implementation is sensitive to systematic uncertainties that arise
%from the need to specify the colors of an unextincted supernova and its use of absolute extinction, neither of which is used in our original model.
\end{abstract}

@@

\subsection{The Two Extrinsic Per-Supernova Parameters $k_0$ and $k_1$}
\color{blue}
The two sets of  $\gamma^i$ parameters  are significantly non-zero. 
None of the 20000 links of 
our Monte Carlo chains for $\gamma$ extend to 0 (see Figures~\ref{global1:fig}--\ref{global5:fig}), we thus claim detection of the
influence of $k_0$ and $k_1$  on supernova magnitudes
with probability $1-5\times 10^{-5}$.
%There is a pair of parameter values,  $k_0^0$ and $k_1^0$, that correspond to supernovae experiencing no extrinsic effect.
%\color{orange}
%For observer filters $X$ and $Y$, we define $A_X(k_0,k_1) =  \gamma^0_X k_0 + \gamma^1_X k_1$, $E(X-Y)=A_X-A_Y$.
%The true extrinsic
%extinction is
%$A^T_X(k_0,k_1)= A_X(k_0,k_1) -  A_X(k^0_0,k^0_1)$.
%The true extrinsic color excess is $E^T(B-V) = A^T_B-A^T_V$, so that the total-to-selective extinction is $R^T_V=A^T_V/E^T(B-V)$.
%Our $\langle k_0 \rangle=0$ and  $\langle k_1 \rangle=0$ constraints
%do not determine $k_0^0$ and $k_1^0$, and so the extrinsic extinction and color excess per supernova are not products of our analysis.
%%our analysis does not specify their values.
%%Recall that there are degeneracies between the intrinsic (Eqn.~\ref{ewsiv:eqn}) and
%%extrinsic (Eqn.~\ref{dust:eqn}) parameters of our model.  Part of that degeneracy is
%%broken by imposing $\langle k_0 \rangle=0$ and  $\langle k_1 \rangle=0$,
%%but this choice does not specify the parameter values.
%Nevertheless, we are able to examine aggregate properties of the extinction.  The extrinsic extinction and color relation can be
%rearranged as
%A_X(k_0,k_1) = R^T_V \left( A_B(k_0,k_1) -  A_V(k_0,k_1)\right) -R^T_V \left(A_B(k^0_0,k^0_1) + A_V(k^0_0,k^0_1)\right)+  A_X(k^0_0,k^0_1) .
%Making the assumption that $R^T_V$ is the same for all supernovae,
%its value can be determined as the slope of the linear relation between $A_X$ and $E(B-V)$.
%Figure~\ref{avebv:fig} shows the relationship between $A_V$ and $E(B-V)$.
%The full set of supernovae does not coincide with the expectations of a single linear relation; rather the supernova population is peaked
%on a ridge line, with a tail distributed below.  That is, for a fixed color excess, the amount of $V$-band extinction is bounded above
%but can have values that extend toward lower values.
%The assumption of a constant $R^T_V$ is not altogether correct; still we perform a linear fit that gives a slope of
%%----- fitz11.py
%$R^T_V=2.77 \pm 0.08$,
%%----
%where the uncertainty is the $68$\% credible interval and does not include model error.  
%%\begin{comment}
%%\color{blue}
%%Our model alone cannot determine $R^T_V = A^T_V/E^T(B-V)$ for individual objects, as $k_0^0$ and $k_1^0$ are unconstrained.
%%For a qualitative view, we choose zeros  such that the $V$-band extinction and color-excess of the bluest supernova roughly correspond to $E(B-V)=0$,
%%$A_V=0$, and such that the ridge line 
%%has close to constant $R^T_V$.
%%\color{black}
%%With this ansatz, we
%%estimate $R^T_V$ per supernova shown in Figure~\ref{avebv:fig}, with the expectations of the
%% \citet{1999PASP..111...63F} (hereafter referred to as \citetalias{1999PASP..111...63F})  model for fixed $R^F_V$ overplotted.
%%With this choice of zeropoints, the ridge line corresponds to $R_V^F\lesssim 2.5$, and even lower values of $R_V^F$ below the ridge line.
%%We emphasize that the shape of the locus in this plot (though not the relative positions of the points)  and the flatness
%%of the ridge line is specific to our choice of zeropoints, and that our model does not specify the zeropoints.
%%A specification of per-supernova $R_V$ requires a model that can identify supernovae that suffer no extrinsic effects,
%%for example by modeling  a constraining distribution for the extrinsic parameters \citep{2016arXiv160904470M}.
%%\color{black}
%   \includegraphics[width=5.2in]{output11/avebv.pdf}
%   %\includegraphics[width=2.8in]{output11/rv.pdf}
%   \caption{
%   $A_V$ versus $E(B-V)$, and the slope $R_V$ of the linear fit to the data.
%   %Right: Shown for qualitative purposes,
%   %$R^T_V=A^T_V/E^T(B-V)$ 
%   %versus $E^T(B-V)$ for the supernova sample, given an ansatz for the definition of an intrinsic supernova. 
%   %Overplotted is the mean and standard deviation of those points in equally divided bins.
%   Overplotted are the expected curves for the \citetalias{1999PASP..111...63F} model
%   for specific choices of $R^F$. Note that these lines are not straight, which is apparent when comparing to the straight black $R_V=2.78$ line.
%   \label{avebv:fig}}
%%\end{comment}
%Until now we have considered $k_0$ and $k_1$ without association with any specific dust model, but they could be interpreted as
%parameters of a single  dust model.
%\color{red}
%Our model for broadband photometry does not specify how spectral features and color parameters affect the supernova spectrum, information
%necessary to predict the magnitude extinction of dust models.
%The lack of correspondence between dust models and broadband extinction is quantified when we apply an $R^F_V=2.5$ SN-frame dust to the
%observed spectra used in our sample, to yield a distribution in the quantity $R_V=A_V/(A_B-A_V)$ with mean and standard deviation of $3.11 \pm 0.35$.
%While an accurate study of fundamental dust properties would incorporate the dust law within the model itself, we now proceed nonetheless
%in making a connection between our model outputs and dust.

We compare our model and its color-coefficient parameters $\gamma^0$ and $\gamma^1$ with the expectations of  dust models.
The 2-parameter linear model
\begin{equation}
A_X = a(X)  A_V + b(X) E(B-V)
\label{f99:eqn}
\end{equation}
approximately describes dust models.
For the case of
$R^F=2.5$ and $A^F_V=0.1$,
and the measurement of the SALT2
\citet{2007A&A...466...11G} $s=1$, $x_1=0$ SN~Ia template at $B$-band peak, the \citetalias{1999PASP..111...63F} model
gives
$a(U,B,V,R,I)=\{0.96,   1.00,   1.00,   0.97,   0.77\}$ and $b(U,B,V,R,I)=\{  1.77,   0.98,   0.12,  -0.50,  -0.53\}$.
The $F$ superscript is added to distinguish \citetalias{1999PASP..111...63F} parameters.
Over the ranges of
 $R^F$ and $A^F_V$,
and the wavelengths under consideration,  the values of the elements of $a$ and $b$ change but only by $<5$\%.

@@

\color{orange}
As a check of the above interpretation, we generate ten simulated datasets with the same numbers and uncertainties as our data, based on a
SALT2 $x_1=0$, $c=0$ template and  \citetalias{1999PASP..111...63F} dust.  The supernovae
are assigned $E^F(B-V)$ parameters equal to the $B-V$ color excesses determined in our sample, and $R^F$ values drawn from a log-normal
distribution with a mean of $\ln{2.5}$ and 0.05 standard deviation.  The $\alpha$, $\beta$, and $\eta$ parameters are set to zero, so that dust and
measurement uncertainty are the only sources of variation in magnitudes.  The input 2.5 is centered
within the 68\% credible interval for $\kappa^{-1}_1$ of 2.38--2.69 from the results of the stacked 10 simulations.
\color{blue}

%
%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=2.95in]{output11_sim/plane0.pdf}
%   \includegraphics[width=2.95in]{output11_sim/plane1.pdf}
%   \caption{
%   Two perspectives of $UVI$- spaces
%of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
%   and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines), for a simulated dataset with
%   input $R^F$ drawn from a log-normal distribution  with a mean of $\ln{2.4}$ and 0.05 standard deviation.
%   The combinations $a(X)+b(X)/2.3$ and  $a(X)+b(X)/2.4$ are shown in the dotted
%   lines, the former is 
%   almost  perfectly  superimposed on  $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$.
%   \label{plane_sim:fig}}
%\end{figure}
%

@@

%
%Another comparison between our model predictions with those of dust is presented in
%Figure~\ref{avebv:fig}, which shows the expectations for measured $A(V)$, $E(B-V)$ for $R^F=1.1, 1.9, 2.5, 3.1$.
%Note that for our broadband measurements, the relationship between $A(V)$ and $E(B-V)$ is subtly
%non-linear.
%Qualitatively, the $R^F \sim 2.4$  most closely matches our linear model fit.
